---
title: "Section 3: Applied Concepts"
author: "Ben LeStage"
format: 
  html:
    embed-resources: true
---

_Example 1_
```{r}
library(tidyverse)
library(palmerpenguins)
theme_set(theme_minimal())

penguins_sum <- penguins |> group_by(species, year) |>
  summarise(n_penguins = n()) |>
  mutate(year = factor(year))

ggplot(data = penguins_sum, aes(x = year, fill = species)) +
  geom_col(aes(y = n_penguins)) +
  theme_minimal() +
  scale_fill_viridis_d()
```

```{r}

ggplot(data = penguins_sum, aes(x = year, y = n_penguins,
                                colour = species, group = species)) +
  geom_line() +
  theme_minimal() +
  scale_colour_viridis_d()
```

The line plot is overall preferable because it is hard to see trends in the species penguins that are not the Gentoo species (which is the baseline).


_Example 2_
```{r}
ggplot(data = penguins, aes(x = species)) +
  geom_bar(fill = "darkslategray4") +
  theme_minimal()
```

```{r}
ggplot(data = penguins, aes(x = species)) +
  geom_bar(fill = "darkslategray4") +
  coord_cartesian(ylim = c(50, 160)) +
  theme_minimal()
```

The first is preferable because if our visualization encodes area or length the y-axis should start at 0.


_Example 3_
```{r}
library(ggbeeswarm)
ggplot(data = penguins, aes(x = species, y = body_mass_g)) +
  geom_beeswarm(alpha = 0.7) +
  theme_minimal()
```

```{r}
ggplot(data = penguins, aes(x = species, y = body_mass_g)) +
  geom_beeswarm(alpha = 0.7) +
  theme_minimal() +
  ylim(c(0, 6500))
```

The first is prefereable ebcause including 0 as a y limit gives this type of graph a lot of wasted space

_Example 4_
pair 1
```{r}
ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point(aes(colour = species)) +
  scale_colour_viridis_d() +
  theme_minimal()
```

```{r}
ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point() +
  facet_wrap( ~ species) +
  theme_minimal() 
```


The first is preferable because it better utilizes the position of points


pair 2
```{r}
penguins <- penguins |> mutate(species_island_sex = interaction(species, 
                                                                island,
                                                                sex))
ggplot(data = penguins |>
         filter(!is.na(sex)), aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point(aes(colour = species_island_sex)) +
  scale_colour_viridis_d() +
  theme_minimal()
```

```{r}
ggplot(data = penguins |> filter(!is.na(sex)), 
                                 aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point() +
  facet_wrap(~ species_island_sex) +
  theme_minimal()
```

The faceted is preferable because too many groups on first, too many colours too hard to distinguish

Improved Plots - 

```{r}
ggplot(data = penguins |> filter(!is.na(sex)), 
                                 aes(x = bill_length_mm, y = bill_depth_mm, colour = sex, shape = species)) +
  geom_point() +
  facet_wrap(~island) +
  theme_minimal()
```

```{r}

ggplot(data = penguins |> filter(!is.na(sex)),
       aes(x = bill_length_mm, y = bill_depth_mm, colour = sex)) +
         geom_point() + 
  facet_grid(species ~ island) + 
  theme_minimal()
```


_Example 5_
```{r}
ggplot(data = penguins, aes(x = island)) +
  geom_bar(fill = "darkslategray4") +
  theme_minimal()
```

```{r}
penguins_island <- penguins |> count(island)
ggplot(penguins_island, aes(x = "", y = n, fill = island)) +
  geom_bar(stat = "identity", width = 1, colour = "white") +
  coord_polar("y", start = 0) +
  theme_void() +
  scale_fill_viridis_d()
```

The bar plot is stronger because it effectively uses position of the bars while the areas of the pie chart are harder to accurately interpret.


_Example 6_

??? In the first, it's easier to distinguish the region while the second has a higher data to ink ratio.














